Investigating shape standardisation using geometric morphometry of pottery from Iron Age northeastern Taiwan

### 2020-04-10 analysis/paper/paper.Rmd
editing results: pca

The elliptic Fourier coefficients were examined by Principal Components Analysis to explore the shape variations across three phases. The first two principal components (PCs) explain `r vari_two`% of the total variances, of which `r pc1`% is explained by the first principle component. With the third component, the first three principal components explain `r vari_three`% of the total variance. According to Figure \@ref(fig:first-two-PCs-byplot), PC1 captures the height of the vessels, from tall to short, and the roundness of the body from round to oval-shaped. PC2 relates to the neck and mouth constriction, from narrow to wide. PC3 explains a smaller portion of the variance (`r pc3`%), which relates to the degree of the flare of the neck, from curved to straight shape. The first two components account for most of the variance in relation to three phases. The results reflect a large overlap in shapes from three occupations phases, especially for shapes in the pre-European and post-European periods. However, the spread of shape distribution indicates a wider variation in shapes in the pre-European and post-European periods compared to those in the Chinese period along both PC1 and PC2 axes. In other words, we find a decrease in shape variance in the Chinese period that is evident in the shorter height and narrower mouth of vessels used in that period.

>>>>

The first two principal components (PCs) of the PCA on the elliptic Fourier coefficients explain `r vari_two`% of the total variance, of which `r pc1`% is explained by the first principle component. With the third component, the first three principal components explain `r vari_three`% of the total variance. PC1 captures the height of the vessels, from tall to short, and the roundness of the body from round to oval-shaped (Figure \@ref(fig:first-two-PCs-byplot)). PC2 relates to the neck and mouth constriction, from narrow to wide. PC3 explains a smaller portion of the variance (`r pc3`%), which relates to the degree of the flare of the neck, from a curved to straight shape. The results reflect a large overlap in shapes from three occupations phases, especially for shapes in the pre-European and post-European periods. However, the spread of shape distribution indicates a wider variation in shapes in the pre-European and post-European periods compared to those in the Chinese period along both PC1 and PC2 axes. In other words, we find a decrease in shape variance in the Chinese period that is evident in the shorter height and narrower mouth of vessels used in that period.
